---
title: 'Korean Drama Analysis'
author: "Nhi Luong and James Le"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: cerulean
  pdf_document: default
subtitle: Fall 2024
urlcolor: blue
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(mosaic)
library(ggthemes)
library(ggplot2)
library(forcats)
library(dplyr)
library(lubridate)

```

## Context and Background

South Korea's movies and TV series have played a crucial role in her
intensive culture export efforts to boost the nation’s ability to “shape
the preferences of others [country] through attraction of values,
policies, and culture as distinct from coercion by economic and/or
military means” (i.e. soft power) [^1]. Its influence in the US
specifically is undeniable as many Korean cultural products have become
a part of the collective mainstream cultural landscape from Kpop music
groups like BTS and Black Pink to the critically and commercially
acclaimed show: Squid Game. The story of Squid Game focuses on the
anxiety and hopelessness of the lower class and neglected groups, whose
only option to escape poverty or paying gambling debt is to participate
in life-staking children's games for the entertainment of the
ultra-wealthy. 

Dive into the same mindset of capitalist horror is the
film Parasite which is the first foreign film to win best picture from
the Oscar and has been considered a historical step for the Korean media
landscape, breaking the old US exclusive-winning convention of this
category. These media provide a medium for many during and after the
pandemics where the same sentiments of the public resonate with the
message of these shows. How do South Korean shows make such an impact?
What can we learn from one of the Four Asian Tigers [^2] cultural
production and publishing’s history and journey? In this project, we
want to explore "Which original network has the highest number of Korean
drama series on weekdays and weekends, and how does it compare to other
days of the week?" and "Which original network has the highest variation
in IMDb ratings, and how does it compare to others in terms of the
median rating?" .

[^1]: Marklund, Carl. “Soft Power.” International Encyclopedia of Human
    Geography, edited by Audrey Kobayashi, 2nd ed., Elsevier Science &
    Technology, 2020. Credo Reference,
    <https://search.credoreference.com/articles/Qm9va0FydGljbGU6NTkwMDQ=?aid=79650>.

[^2]: Day, D.-C. “Four Asian Tigers’ Political and Economic Development
    Revisited 1998-2017: From the Perspective of National Identity”.
    Asian Journal of Interdisciplinary Research, vol. 4, no. 4, Dec.
    2021, pp. 54-61, <doi:10.54392/ajir2147>.

```{r, echo=FALSE}
kdrama_tidy <- read_csv("~/Downloads/kdrama_tidy.csv")
```

### IMDb Ratings Across Original Networks

But how popular is the Korean Wave with its cultural products of television that have taken the world by storm and especially in the West? We would like to touch the surface of the Kdrama popularity through the IMDb rating of these shows while comparing them between networks to assess such impact. Here, we define original network of a show as a broadcasting channel that invests in the show and airs it first. 

```{r, echo=FALSE, fig.cap="Figure 1: IMDb Ratings distribution by Original Networks for Korean Dramas.This box plot illustrates the distribution of IMDb ratings for Korean dramas produced by 20 original networks. The data reveal that most networks have median ratings between 6 and 8. Netflix stands out with the highest median rating, while Naver TV Cast has the lowest. Outliers are present across several networks, suggesting significant variability in ratings for some shows. The data were collected from mydramalist.com and include IMDb ratings for 810 dramas spanning from 2002 to 2022. The figure provides a descriptive overview of users rating.", fig.alt="This is a box plot showing the IMDb ratings distribution for dramas from 21 Korean original networks. The x-axis represents IMDb ratings ranging from 0 to 10, and the y-axis lists the networks ordered by their median IMDb ratings with Netflix at the top and Naver TV Cast at the bottom. Most shows across all networks have ratings between 6 and 8. Netflix exhibits the highest median rating, while Naver TV Cast has the lowest median rating. Outliers are visible for several networks, indicating a few shows with exceptionally high or low ratings."}

#Graph 1: IMDb ratings

kdrama_tidy |>
  mutate(original_network = fct_collapse(original_network,
                                         "SBS" = c("SBS", "SBS Plus"),
                                         "MBC" = c("MBC", "MBC every1", "MBC Dramanet"), 
                                        "Others" = c("vLive", "Tooniverse", "Sohu TV", "BS Japan", "Amazon Prime"), "tvN" = c("tvN", "XtvN")),
        original_network = fct_reorder(original_network, imdb_rating, .na_rm = TRUE)) |>
  ggplot(aes(x = original_network, y = imdb_rating)) +
  geom_boxplot() +
  labs(
    title = "IMDb Ratings Distribution by Original Network",
    y = "IMDb Ratings",
    x = "Original Network"
  ) +
  theme_minimal() +
  coord_flip()
```

__Observation__

-   We explore the IMDb ratings distribution across 21 Korean original
    networks and see that Netflix has the highest median IMDb ratings,
    surpassing national public networks such as KBS, SBS, and MBC by
    half a point. Among these original networks, Netflix is one of a few
    original network that is a streaming service, but it is doing very
    well. As Netflix is a global streaming service, this suggests
    Netflix's original Korean dramas are doing very well and receiving a
    lot of attention from global audience. This relates to the concept
    of soft power we discuss in the introduction where Korean culture
    and values are getting recognition from the international level.

### Aired Day of K-drama Series 

One way to examine the strategy to their success would be through the classic strategy of TV networks in their choice of aired day to be competitive with other networks. We will visualized that concept through the top 10 original networks below.

```{r, echo=FALSE, results='hide'}
kdrama_tidy |>
  group_by(original_network) |>
  summarize(n = n()) |>
  slice_max(n, n = 10)

```


```{r, echo=FALSE, fig.cap="Figure 2: Number of Korean Drama Series by Days of the Week Across Top 10 Original Networks. This chart shows that KBS2, SBS, and tvN aired the highest number of Korean drama series, with peaks on Monday, Tuesday, Wednesday, Thursday, and Saturday. Data were collected from the publicly available website mydramalist.com, where users contribute information about Asian TV shows and movies. The dataset includes 810 distinct shows spanning from 2002 to 2022. No statistical tests were performed.", fig.alt = "This is a faceted bar chart showing the distribution of Korean dramas by the day of the week they aired, categorized by their original networks. The x-axis represents the number of dramas released, ranging from 0 to 3000, while the y-axis lists from top to bottom Monday to Sunday of a week.. Each facet corresponds to the top 10 networks in order from top to bottom, left to right: jTBC, KBS1, KBS2, MBC, Naver TV Cast, Netflix, OCN, SBS and Viki. Networks like KBS2, SBS, and tvN dominate with the highest number of shows aired across the week, especially on Monday, Tuesday, Wednesday, Thursday, and Saturday.", fig.width=6, fig.height=6}

#Graph 2: Days of Week

filtered_data <- kdrama_tidy |>
  mutate(aired_on = factor(aired_on, 
                           levels = rev(c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")),
                           labels = rev(c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))) |>
  filter(original_network %in% c("tvN", "SBS", "KBS1", "KBS2", "MBC", "jTBC", "Netflix", "Viki", "OCN", "Naver TV Cast"))

filtered_data |>
    ggplot(aes(y = aired_on, fill = aired_on)) +
    geom_bar(aes(x = after_stat(count)), stat = "count", show.legend = FALSE) +
    facet_wrap(~original_network) +
    labs(x = "Number of Korean Drama Series",
         y = "Day of the Week",
         title = "Number of Korean Drama Series by Day of the Week \nAcross Original Networks",
         subtitle = "Top 10 Original Networks") +
  theme_bw() +
  scale_fill_colorblind()
```

__Observation__

-   This graph illustrates that Tuesday and Wednesday are the most
    popular days for airing dramas across multiple networks. Conversely,
    Sundays and Fridays show relatively lower activity in terms of drama
    releases, indicating a preference for mid-week broadcasting. One
    possible explanation for these trends due to market demands. Sunday
    is the final weekend day which leads to lower viewership while the
    reason Friday has less demand might be due to overtime
    working/studying schedule or the population at large going out and
    not staying at home watching TV. Another notable observation would
    be that the top four networks dominated the market where KBS, MBC
    and SBS are public television networks and tvN is a private one.
    Each network also have their own airing strategies as tvN aired the
    most shows on Saturday, KBS on Monday and Tuesday while MBC have the
    highest proportion on Wednesday and Thursday. We can also observed that investment in the public sector (KBS, SBS) is as important as the private sector (tvN) for the television industry as the government heavily subsided the industry through investment in film education from the Korean Film Academy, interest-free or low-interest loans for enterprise [^4] and tax policies[^5] that encourage local business to invest in film production beside their own direct investment to the industry [^6].
    
    
[^4]: Kang, Jennifer M. “Better than Television: The Rise of Korean Web Dramas.” International Journal of Cultural Studies, vol. 24, no. 6, May 2021, p. 136787792110145, https://doi.org/10.1177/13678779211014532.

[^5]: PricewaterhouseCoopers International Limited. “Korea, Republic of - Corporate - Tax Credits and Incentives.” Pwc.com, 2024, taxsummaries.pwc.com/republic-of-korea/corporate/tax-credits-and-incentives. Accessed 12 Dec. 2024.

[^6]: Shim, Doobo. “Whiter the Korean Flim Industry?” Acta Koreana, vol. 14, no. 1, June 2011, pp. 213–27, https://doi.org/10.18399/acta.2011.14.1.010. Accessed 1 Apr. 2020.

    
### K-drama on Netflix Before and After Covid

On March 11, 2020, WHO declared COVID-19 outbreak a global pandemic [^3] that continued to last for the next 3 years. Following the announcement was a string of global and national shutdowns. Because of that, more people spent more time inside watching shows on Netflix and got to learn more about Korean drama. In this section, we would like to look at Korean drama series on Netflix to identify changes in the number of drama series as well as the most popular genres among the series before and after the Covid pandemic. We have decided to use "2020" as a mark for before and after Covid pandemic instead of "March 2020" for simplicity. 


[^3]: Cucinotta, Domenico, and Maurizio Vanelli. “WHO Declares COVID-19 a Pandemic.” Acta Bio Medica: Atenei Parmensis, vol. 91, no. 1, 2020, pp. 157–60. PubMed Central, https://doi.org/10.23750/abm.v91i1.9397.

```{r, echo=FALSE, results='hide'}
kdrama_tidy |>
  filter(original_network == "Netflix", platforms == "Netflix") |>
  mutate(start_airing = mdy(start_airing),
    start_year = year(start_airing),
    start_year = as.factor(start_year),
    start_year = fct_collapse(start_year, "Before 2020" = c("2016", "2017", "2018", "2019"),
                              "After 2020" = c("2020", "2021", "2022"))) |>
  distinct(drama_name, .keep_all = TRUE) |>
  count(start_year)
```


```{r, echo=FALSE, fig.alt="A bar plot comparing counts of Korean drama shows Before 2020 and After 2020. The x axis represents the year categories with Before 2020 on the left and After 2020 on the right, while the y axis represents the number of shows, ranging from 0 to 2500 shows. The bar for Before 2020 has a height of 1170, meaning there are 1170 Korean drama series on Netflix before 2020, marked by a dashed red horizontal line. The bar for After 2020 has a height of 2382, meaning there are 2382 Korean drama series on Netflix after 2020, marked by a dashed blue horizontal line. The y axis is labeled with increments of 500, including custom labels at 1170 and 2382 for emphasis. The plot title reads Comparison of Counts Before and After 2020. We see that the number Korean shows on Netflix after 2020 is twice as much as the number Before 2020", fig.cap="Figure 3: The number of Korean drama series on Netflix after 2020 is two times greater than the number of Korean shows on Netflix before 2020. The data were collected from mydramalist.com and include IMDb ratings for 810 dramas spanning from 2002 to 2022. The years were grouped into two categorical groups for comparison. No statistical tests were performed.", fig.width=6, fig.height=6}

# Graph 3a: Netflix shows before and after Covid

kdrama_tidy |>
  filter(original_network == "Netflix", platforms == "Netflix") |>
  mutate(start_airing = mdy(start_airing),
    start_year = year(start_airing),
    start_year = as.factor(start_year),
    start_year = fct_collapse(start_year, "Before 2020" = c("2016", "2017", "2018", "2019"),
                              "After 2020" = c("2020", "2021", "2022"))) |> 
  distinct(drama_name, .keep_all = TRUE) |>
  ggplot(aes(x = start_year)) +
  geom_bar(fill = "steelblue") +
  geom_hline(yintercept = 27, color = "red", linetype = "dashed") + 
  geom_hline(yintercept = 63, color = "blue", linetype = "dashed") + 
  scale_y_continuous(breaks = c(seq(0, 100, by = 10), 27, 63)) +
  labs(
    x = "Year",
    y = "Number of Shows",
    title = "Comparison of Counts Before and After 2020"
  ) +
  theme_minimal()
```

* We briefly look at the number of Korean drama shows on Netflix before 2020 and after 2020 and see that after the announcement of the COVID-19 a global pandemic up until 2022, Netflix has added 36 more Korean drama shows into their collection, showing Netflix's interest in investing into Korean shows in these recent years. Naturally, our next question would be "Then does Netflix have a preference of Korean drama shows' genres that they want to add that can attract more viewers?" We look at this question in the following graph. 

```{r, fig.width=8, fig.height=6, echo=FALSE, fig.alt="A line graph titled Top 5 Korean Drama Genres on Netflix displays the number of shows with at least one of the top 5 genres from 2016 to 2022. The y axis represents Number of Shows ranging from 0 to 160, and the x axis represents Year Released ranging from 2016 to 2022 with a one year increment. The chart includes five colored lines, each representing a genre. Romance is black, action is orange, mystery is blue, thriller is green, and comedy is yellow. A vertical red line marks the year 2020 to compare the change in Korean drama genres on Netflix before and after Covid. A legend to the right identifies each genre by color. From the graph, we see genre romance shows significant growth over the years, peaking around 2020, dropping significantly in 2021, but climbing back in 2022. On the other hand, shows with mystery and thriller genres have a huge jump in number of shows between 2020 and 2021.", fig.cap="Figure 3: Top 5 Korean Drama Genres on Netflix from 2016 to 2022. The red vertical line marks the transition year 2020. Shows with omance genres exhibited the fastest growth, peaking at around 155 in 2020. Data were sourced from the publicly available website mydramalist.com, where users contribute information about Asian TV shows and movies. No statistical tests were conducted."}

# Graph 3b: Top 5 Korean Drama Genres on Netflix from 2016 to 2022

netflix_top5_genres <- kdrama_tidy |>
  distinct(drama_name, .keep_all = TRUE) |>
  count(genres) |>
  slice_max(n, n = 5)

kdrama_tidy |>
  filter(original_network == "Netflix", platforms == "Netflix") |>
  mutate(start_airing = mdy(start_airing),
    start_year = year(start_airing)) |>
  group_by(genres, start_year) |>
  summarize(count = n(), .groups = "drop" ) |>
  semi_join(netflix_top5_genres, by = "genres") |>
  mutate(genres = fct_reorder2(genres, start_year, count)) |>
  ggplot(aes(x = start_year, y = count, color = genres)) +
  scale_color_colorblind ()+
  geom_point() +
  geom_line() +
  labs(
    title = "Top 5 Korean Drama Genres on Netflix",
    subtitle = "From 2016 to 2022",
    color = "Genres",
    x = "Year Released",
    y = "Number of Shows"
  ) +
  scale_fill_colorblind() +
  geom_vline(xintercept = 2020, col = "red") +
  theme_minimal()

```

__Observation__

- This graph indicates a shift in focus toward action and romance genres while reducing investment in thriller and comedy content from Netflix in Kdrama. We can see that romance dramas experienced a sharp increase, peaking around 2020, followed by a slight decline but maintaining it’s top position even in 2022. The action genres initially grew at a slower pace but also grew significantly after 2020. Mystery and thriller genres followed similar trajectories, with significant activity between 2018 and 2020 but a noticeable decline after 2021. Comedy dramas showed a slower growing trend which peaked in 2019 and then declined gradually. The red vertical line marking 2020 reflects the impact of a pandemic which might change production priorities and viewer preferences. 

## Conclusion

Through the project, we have looked at IMDb ratings of Korean shows by original networks, their air day in the week, and changes of Korean shows on Netflix. We see that overall, the Korean drama industry is investing a lot into their shows so that the shows can be known more globally. As a result of their strategy and popularity of Korean drama, these shows receive more investment from Netflix and appear on the streaming service more, which makes a big profit for Netflix while as the same time receive worldwide popularity. We believe that Korean drama shows will continue to thrive even more in the future. 
